---
title: NREL Load Shapes
author: Chris Cote
date: '2020-03-05'
slug: nrel-load-shapes
categories: []
tags: []
---

## This post is a quick look at new data from NREL: "Electrification Futures Study Load Profiles" 

The website describes the datasets:

This data set includes hourly load profiles developed for the Electrification Futures Study (EFS). The load profiles represent projected end-use electricity demand for various scenarios of electrification (Reference, Medium, High) and technology advancement (Slow, Moderate, Rapid), developed using the EnergyPATHWAYS (EP) model, as described in Mai et al. (2018). The EP load profiles were further calibrated for use in the ReEDS capacity expansion model (detailed in forthcoming EFS analyses), thus numerical values in this data set differ slightly from those reported in Mai et al. (2018). Load profiles are provided for a subset of years (2018, 2020, 2024, 2030, 2040, 2050) and are aggregated to the state, sector, and select subsector level.

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
#data dl from https://data.nrel.gov/submissions/126?utm_source=NREL+Electrification+Futures+Study 

#had to unzip with terminal on mac, note on website says:
#NOTE: Due to the file size, Mac users may experience issues decompressing the zip files using the Mac Archive Utility. In those cases, decompressing using the command line is recommended.
#just "unzip filename"

```

```{r message=FALSE, warning=FALSE}
efs <- data.table::fread("~/Downloads/EFSLoadProfile_Reference_Moderate.csv")
```

How does demand for electric-charging vehicles change by decade over the first 100 hours of each year... (not so helpful but fun looking)
```{r}
efs %>% as_tibble() %>% 
  filter(State == "MA") %>% 
  filter(Sector == "Transportation") %>% 
  filter(Subsector == "light-duty vehicles") %>% 
  # filter(Year == 2030) %>% 
  filter(LocalHourID < 100) %>% 
  mutate(sector_subsector = paste0(Sector, " ", Subsector)) %>% 
  ggplot() + geom_line(aes(LocalHourID, LoadMW, color = factor(Year))) +
  facet_wrap(~sector_subsector) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```



```{r}
annual_growth <- efs %>% as_tibble() %>% 
  filter(Sector == "Transportation") %>% 
  filter(Subsector == "light-duty vehicles") %>% 
  group_by(Year, State) %>% 
  summarise(LoadMW = sum(LoadMW, na.rm = TRUE)) %>% 
  group_by(State) %>% 
  mutate(growth = (LoadMW-lag(LoadMW))/lag(LoadMW)) %>% 
  ungroup() 
  
annual_growth %>% 
  filter(State %in% c("MA", "NY", "NJ", "TX", "CA", "IL")) %>% 
  ggplot() + geom_line(aes(Year, growth, color = State)) + 
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```


```{r}
ma_sector <- efs %>% as_tibble() %>% 
  filter(State == "MA") %>% 
  group_by(Sector, Year, LocalHourID) %>% 
  summarise(LoadMW = sum(LoadMW)) %>% 
  ungroup()
  
ma_sector
```

The commercial sector has the largest seasonal peak and uses the most electricity overall.
```{r}
ma_sector %>% 
  group_by(Year, Sector) %>% 
    mutate(day = rep(1:(length(unique(ma_sector$LocalHourID))/24), each = 24)) %>% 
  group_by(day, Sector, Year) %>% 
  summarise(LoadMW = mean(LoadMW)) %>% 
  ungroup() %>% 
  ggplot() + geom_line(aes(day, LoadMW, color = factor(Year))) + 
  facet_wrap(~Sector) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```
From a high level view, it looks like their estimates have commercial use increasingly significant in 2050 while the residential load stays put. Transportation in MA doesn't grow to significant levels, meaning EVs (in this moderate reference case), never really have a big impact. Industrial is steady.

What's driving commercial use? 

```{r}
efs %>% 
  as_tibble() %>% 
  filter(State == "MA", 
         Sector == "Commercial") %>% 
  group_by(Year, Subsector) %>% 
  summarise(LoadMW = mean(LoadMW)) %>% 
  ungroup() %>% 
  ggplot() + geom_col(aes(Subsector, LoadMW, fill = factor(Year)), position = "dodge") +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```

How helpful... an enormous jump in "other" in 2050. Maybe the EnergyPathways docs say more. For another day. 


Does the variability of demand in MA change over time? I suppose there are several ways to measure this, but I'll take the daily range (max - min). Can we see anything in a daily plot?

```{r}
ma_sector_hour <- ma_sector %>%
  group_by(Year, LocalHourID) %>% 
  summarise(LoadMW = sum(LoadMW)) %>% 
  group_by(Year) %>% 
  mutate(day = rep(1:(length(unique(ma_sector$LocalHourID))/24), each = 24)) %>% 
  group_by(day, Year) %>% 
  mutate(hour_day = 1:n()) 
  
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  ungroup() %>% 
  ggplot() + geom_line(aes(day, daily_range, color = factor(Year))) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
  
  
```

Not much variation to see from this point of view! 

Let's try it in an annual plot

```{r}
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  group_by(Year) %>% 
  summarise(daily_range = mean(daily_range)) %>% #this part changes
  ggplot() + geom_col(aes(Year, daily_range, fill = factor(Year))) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```

Or since what we're really trying to do here is find the standard deviation, let's just check that per day
```{r}
ma_sector_hour %>% 
  summarise(sd = sd(LoadMW)) %>% 
  group_by(Year) %>% 
  summarise(sd = mean(sd)) %>% #this part changes
  ggplot() + geom_col(aes(Year, sd, fill = factor(Year))) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```
Same plot, just with different units.


Do the load peaks and nadirs stay the same over time? 
```{r}
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("max" = "LoadMW", "Year", "day")) %>% 
  rename(max_hour = hour_day) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("min" = "LoadMW", "Year", "day")) %>% 
  rename(min_hour = hour_day) %>% 
  ungroup() %>% 
  ggplot() + geom_ribbon(aes(day, ymin = min_hour, ymax = max_hour), 
                         fill = "lightblue", alpha = 3/4) + 
  facet_wrap(~Year) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))

```

Ok this is pretty unhelpful. We can see how the load shape changes over the year, but it's pretty difficult to see whether there's any variation across years.

Let's look at just one year, since we're here.

```{r}
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("max" = "LoadMW", "Year", "day")) %>% 
  rename(max_hour = hour_day) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("min" = "LoadMW", "Year", "day")) %>% 
  rename(min_hour = hour_day) %>% 
  ungroup() %>% 
  filter(Year == 2018) %>% 
  ggplot() + geom_ribbon(aes(day, ymin = min_hour, ymax = max_hour), 
                         fill = "lightblue", alpha = 3/4) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```


Let's try another way of looking at this...
```{r}
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("max" = "LoadMW", "Year", "day")) %>% 
  rename(max_hour = hour_day) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("min" = "LoadMW", "Year", "day")) %>% 
  rename(min_hour = hour_day) %>% 
  ungroup() %>% 
  mutate(yday = day + lubridate::as_date(paste0(Year,"-01-01"))-1,
         month = lubridate::month(yday)) %>% 
  group_by(Year, month) %>% 
  summarise(max_hour = mean(max_hour),
            min_hour = mean(min_hour)) %>% 
  ungroup() %>% 
  ggplot() + geom_segment(aes(Year, xend = Year, y = min_hour, yend = max_hour)) + 
  geom_point(aes(Year, max_hour)) +
  geom_point(aes(Year, min_hour)) + 
  facet_wrap(~month) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"))
```
This is still only so-so but the point is the hours when the max and mins hit don't change much over the decades




```{r}
ma_sector_hour %>% 
  summarise(max = max(LoadMW),
            min = min(LoadMW), 
            daily_range = (max - min)) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("max" = "LoadMW", "Year", "day")) %>% 
  rename(max_hour = hour_day) %>% 
  left_join(ma_sector_hour %>% 
              select(hour_day, LoadMW), by = c("min" = "LoadMW", "Year", "day")) %>% 
  rename(min_hour = hour_day) %>% 
  group_by(Year) %>% 
  mutate(yday = day + lubridate::as_date(paste0(Year,"-01-01"))-1,
         month = lubridate::month(yday)) %>% 
  # filter(Year == 2018) %>% 
  ungroup() %>% 
  ggplot() + 
  # geom_point(aes(max, max_hour, color = factor(Year)), alpha = 2/3) + 
  geom_point(aes(max_hour, max, color = factor(Year)), shape = 21) + #outline
  facet_wrap(~month) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_line(color = "gray90"),
        axis.text.x  = element_text(angle = 90))
```

Not sure why there is so much variation by day in the summer months! Any ideas? 
