---
title: "Grandpa in Law in WWII"
author: "Chris Cote"
date: '2020-02-22'
output:
  html_document: 
    fig_height: 6
    fig_width: 9
slug: john-ross-in-wwii
tags: []
categories: spatial
---
My wife's paternal grandfather took meticulous notes throughout his life, including during his service in the Army during World War II. This Christmas we found and looked through some of those notes, and I figured I could turn then into a fun dataset and map.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)

```

create point data frame
```{r message=FALSE, warning=FALSE}

jr_points <- read_csv("~/Documents/HKS/rtutorial/data/johnross/jr_raw.csv") %>% 
  mutate(year = str_extract(`Date Arrived`, "-4[0-9]") %>% str_replace(., "-", ""),
         month = str_extract(`Date Arrived`, "[A-Z][a-z][a-z]") %>% str_replace_all(., "-", ""),
         day = str_extract(`Date Arrived`, "^[0-9]([0-9])?"),
         year = ifelse(is.na(year), lag(year), year),
         date = paste0("19",year,"-", month,"-", day) %>% as.Date("%Y-%b-%d")) %>% 
  filter(!is.na(Lat)) %>%
  select(date, year, month, day, everything()) %>% 
  mutate(index = 1:n()) %>% 
  set_names(str_replace(str_to_lower(names(.)), " ", "_")) %>% 
  mutate(date_arrived = as.Date(paste0("19", year, "-", month, "-", day), format = "%Y-%b-%d")) %>% 
  arrange(date_arrived) %>% 
  mutate(date_left = lead(date_arrived)) %>% 
  mutate(location = paste(location_1, location_2, location_3, sep = ", ")) %>% 
  select(date_arrived, date_left, location, what, lat, lon) %>% 
  mutate(new_location = ifelse(location == lag(location), 0 , 1),
         new_location = ifelse(row_number() == 1, 0, new_location),
         new_location = cumsum(new_location)) %>% 
  group_by(new_location) %>% 
  fill(what) %>% 
  ungroup() %>% 
  distinct(new_location, .keep_all = TRUE) %>% 
  mutate(date_left = lead(date_arrived)) %>% 
  mutate(country = ifelse(str_detect(location, "Italy"), "Italy", 
                          ifelse(str_detect(location, "USA"), "USA", "NA")),
         lead_country = lead(country),
         lead_country = ifelse(is.na(date_left), "USA", lead_country),
         map = paste0(country, lead_country), 
         map2 = ifelse(map == lag(map), 0, 1),
         map2 = ifelse(is.na(map2), 1, map2)) %>%  
  group_by(map) %>% 
  mutate(map2 = cumsum(map2)) %>% 
  ungroup() %>% 
  mutate(map = paste0(map,map2))

```

```{r}
jr_points
```


```{r}
jr_labels <- 
   jr_points %>% distinct(date_arrived, date_left, location, lat, lon, map) %>% 
   mutate(label = paste0(location, ": ",  scales::date_format("%b-%d-%y")(date_arrived), " to ", scales::date_format("%b-%d-%y")(date_left))) %>% 
   mutate(label = str_replace(label, ", NA", "")) %>% 
   mutate(country = ifelse(str_detect(location, "Italy"), "Italy", 
                           ifelse(str_detect(location, "USA"), "USA", "NA")),
          lead_country = lead(country),
          lead_country = ifelse(is.na(date_left), "USA", lead_country)) %>% 
   mutate(duration = as.numeric(date_left - date_arrived)) %>% 
   group_by(location) %>% 
   mutate(duration = sum(duration, na.rm = TRUE)) %>% 
   ungroup()

jr_labels
```

Create long function to make map plot
```{r}
make_map <- function(choose_map) {
 jr_labels <- 
   jr_points %>% distinct(date_arrived, date_left, location, lat, lon, map) %>% 
   mutate(label = paste0(location, ": ",  scales::date_format("%b-%d-%y")(date_arrived), " to ", scales::date_format("%b-%d-%y")(date_left))) %>% 
   mutate(label = str_replace(label, ", NA", "")) %>% 
   mutate(country = ifelse(str_detect(location, "Italy"), "Italy", 
                           ifelse(str_detect(location, "USA"), "USA", "NA")),
          lead_country = lead(country),
          lead_country = ifelse(is.na(date_left), "USA", lead_country),
          map = factor(paste0(country, lead_country))) %>% 
   mutate(duration = as.numeric(date_left - date_arrived)) %>% 
   group_by(location) %>% 
   mutate(duration = sum(duration, na.rm = TRUE)) %>% 
   ungroup() %>% 
   mutate(duration = (scale(duration) + 3)^5)
  if(missing(choose_map)){
    
    choose_map <-""

    
  }else{
    jr_points <- jr_points %>% 
    filter(str_detect(map, choose_map))
  jr_labels <- jr_labels %>% 
    filter(str_detect(map, choose_map)) }
 

  
  if(!missing(choose_map) & choose_map == "USAUSA"){
    map_usa <- ggplot2::map_data("state")
ggplot(jr_points) + 
    geom_map(data = map_usa, map = map_usa, aes(map_id = region), 
             fill = "gray90", alpha = 1/2,
             color = "gray50") + 
    geom_curve(data = jr_points %>% 
                 mutate(alternate = rep_along(date_arrived, 1:2),
                        alternate = factor(alternate)),
               aes(x = lon, xend = lead(lon), y = lat, yend = lead(lat), lty = alternate),
               arrow = arrow(length = unit(0.2, "inches"), type = "closed"), show.legend = FALSE) +
    geom_point(data = jr_labels,  
               aes(x = lon, y = lat, color = location, size =  duration), show.legend = FALSE) +
    ggrepel::geom_label_repel(data = jr_labels %>% 
                               mutate(number = row_number(),
                                      number_location = paste0(number, ": ", str_replace_all(label, ", USA|, Italy", ""))) %>% 
                               group_by(location) %>% 
                               mutate(number_location = paste(number_location, collapse = "\\\n")) %>% 
                               distinct(number_location, lat, lon), 
                             aes(x = lon, y = lat, label = number_location), size = 2.7, show.legend = FALSE) +
    coord_equal() + 
    theme_void()
  }else {
    world <- ggplot2::map_data("world")
    ggplot(jr_points) + 
      geom_map(data = world, map = world, aes(map_id = region), 
               fill = "gray90", alpha = 1/2,
               color = "gray50") + 
      geom_curve(data = jr_points %>% 
                   mutate(alternate = rep_along(date_arrived, 1:2),
                          alternate = factor(alternate)),
                 aes(x = lon, xend = lead(lon), y = lat, yend = lead(lat), lty = alternate),
                 arrow = arrow(angle = 120, length = unit(0.1, "inches"), type = "closed"), show.legend = FALSE) +
      geom_point(data = jr_labels, 
                 aes(x = lon, y = lat, color = location, size = log(duration)),  show.legend = FALSE) +
      ggrepel::geom_label_repel(data = jr_labels %>%
                                 mutate(number = row_number(),
                                        number_location = paste0(number, ": ", str_replace_all(label, ", USA|, Italy", ""))) %>%
                                 group_by(location) %>%
                                 mutate(number_location = paste(number_location, collapse = "\\\n")) %>%
                                 distinct(number_location, lat, lon),
                               aes(x = lon, y = lat, label = number_location), size = 3, show.legend = FALSE) +
      coord_equal() + 
      theme_void() +
      theme(panel.background = element_rect(fill = "NA")) 
  } 
  
}

```


use the function!

These two maps show where he went during the entire length of his service in the Army, which took him across the US and Italy, with a vacation into Switzerland (not shown).


Service in USA  -- lots of travel
```{r fig.align="center", fig.height=6, fig.width=9, message=FALSE, warning=FALSE}

make_map("USAUSA")

```


Service in Italy. This was harder to map because he continued to pass through the same cities again and again. 

```{r fig.align="center", fig.height=6, fig.width=9, message=FALSE, warning=FALSE}

make_map("ItalyItaly")

```

I would have liked to use geom_sf with this, but geom_curve was more flexible! 

These aren't perfect maps, I hope to keep iterating, eventually playing with rasters, improving the arrows, and understanding if there are better methods of adding labels. Though I did think adding the repeat trips inside the same label with paste(collapse) was clever!
